---
title: "NBA Salaries"
author: "Jeremy Hansen"
date: "1/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Started ##

I found a very helpful [guide](https://statisticaloddsandends.wordpress.com/2020/08/29/nba-salaries/) to play around with NBA salary data.  The charts that I have made are only going to deal with the data from 2000-2017.  

The first thing to do is load all the packages that we will use.  I used tidyverse, plyr, dplyr, utils, ggplot2, ggthemes, readxl, and plotly.  The important thing here is that you load plyr before dplyr, otherwise you will run into problems later.  

The data is downloaded from [here](https://data.world/datadavis/nba-salaries).

```{r, echo = FALSE, include = FALSE}
library(tidyverse)
library(plyr)
library(dplyr)
library(utils)
library(ggplot2)
library(ggthemes)
library(readxl)
library(plotly)
```
```{r, message=FALSE, warning = FALSE}

players <- read_csv("players.csv") %>% 
  select(`_id`, name)
salaries <- read_csv("salaries_1985to2018.csv") %>%
  inner_join(players, by = c("player_id" = "_id"))
```

## Some data manipulation and basic chart ##

Now that the data is here, lets take a look at it.

```{r}
head(salaries)
```

Because I am only interested in 2000-2017, and only some of the columns, we'll handpick the ones that matter. I chose to multiply total salary by .000000001, in order to make the scale be in billions.

```{r, warning=FALSE}
salaries1 <- salaries %>% 
  filter(season_start >= 2000) %>%
  select(player_id, name, salary, year = season_start, team)

# total salary by year
yearly <- salaries1 %>% 
  group_by(year) %>% 
  dplyr::summarise(tot_salary = sum(salary) * .000000001)

#just a rough graph
ggplot(yearly, aes(x = year, y = tot_salary)) +
  geom_point(color = "cornflowerblue") +
  geom_line(color = "cornflowerblue") +
  expand_limits(y = .5) +
  labs(x = "Year", 
       y = "Total League Salary (Billions)", 
       title = "Total Salary of the League by Year") +
  theme_few()
```

We see that the total salary has been increasing each year, except from about 2009-2014.  More recently, the amount of salary has been increasing more exponentially.  This brought up a question: How has the increase in total salary compared to yearly inflation rate?

I went to [FRED](https://fred.stlouisfed.org/?gclid=CjwKCAiAl4WABhAJEiwATUnEF4fuIbnSi4txBwJuFGlgGRSrvhM6qq2W2GPuXVSaAu5WZ65FBnx1nBoCJigQAvD_BwE) and looked up the inflation rates for 2000-2017.  


```{r}
rates <- c(1.2, 1.8, 2.4, 2.1, 1.3, 0.1, 1.6, 1.5, 2.1, 3.2, 1.6, -0.4, 3.8, 2.8, 3.2, 3.4, 2.7, 2.3, 1.6, 2.8, 3.4)
```







```{r}

inflation_df <- data.frame(year = 2000:2017,


gg <- ggplot(data = team_salaries, aes(year, tot_salary)) +
  geom_line(aes(group = team), color = "gray70") +
  expand_limits(y = 0) +
  geom_smooth(se = F, color = "salmon") +
  labs(title = "team salaries over time", 
       subtitle = "(each line is a team, pink is the average)",
       x = "year",
       y = "total team salary (millions)") +
  theme_few()

ggplotly(gg)


```